### 2024-10-06 AISTAT25/aistats.tex
Yoavs final

@@

\begin{abstract}%
  We present a general-purpose active learning scheme that maintains a
  collection of balls of different sizes and uses label
  queries to identify those with a strong bias towards one particular
  label. When two such balls intersect and have different
  labels, the region of overlap is treated as a ``known unknown'' and
  is targeted in future active queries. We give label complexity
  bounds for this method that do not rely on assumptions about the
  data, and we instantiate them in several cases of interest.
\end{abstract}

@@

We begin with a large collection of \emph{balls}, $\B$, of varying
sizes (Fig.~\ref{fig:challenges}(a)). We use label queries to assign
\emph{biases} $\yh(B) \in \{-1,0,+1\}$ to balls $B \in \B$, beginning
with the largest balls. A bias of $+1$ means that the average of 
$\eta(x)$ over $x \in B$ is significantly positive; bias $-1$ means it is
significantly negative; and bias $0$ means the average $\eta$ is close
to 0. As time goes on, we get biases for progressively smaller balls,
as needed. The final label assigned to any point $x$ is based on the
biases of balls containing $x$.

@@

The rest of the paper is organized as
follows. Section~\ref{sec:relatedwork} describes related work.
Section~\ref{sec:algorithm} presents the active learning algorithm and
Section~\ref{sec:discrete} analyzes it in the finite population
setting, taking $X$ to be an arbitrary point-set and allowing any
$\eta$ function. We identify two \emph{critical levels} for any
$x \in X$: two scales (of ball sizes) that control how many queries
are sufficient for $x$ to be correctly labeled
(Theorem~\ref{thm:label-complexity}). We instantiate these bounds in
canonical one-dimensional settings (Theorems~\ref{thm:oned-massart}
and \ref{thm:oned-monotonic}) to get label complexities logarithmic in
$|X|$. In section~\ref{sec:experiment} we describe an experiment that supports the theoretical bound given in Theorem~\ref{thm:oned-massart}.

@@

\section{Definitions}

The two basic entities in our algorithm are the collection of points
$X$ and a collection of balls $\B$. Sampling is organized around $\B$:
these balls are the atomic sets on which we assess label
bias.

@@

We now describe the simplified algorithm, here referred to simply as
the algorithm. The algorithm loops over increasing levels.  At each
level it identifies the region of ``known unknowns'' which approximates
the boundary. The algorithm then makes queries in the vicinity of the
boundary to refine it and proceeds to the next finer level.

What follows is a detailed explanation of  Figure~\ref{alg:simple}.
The outer loop of the algorithm iterates over
levels, starting with $\ell=0$ that corresponds to balls with at least half of the points in $X$, then moving onto $\ell=1$ which has balls with between $1/4$ and $1/2$ of the points in $X$, and so on.
%and ending with $\ell=\log |X|-1$,
%corresponding to balls that contain a single point. %Initially, all balls are labeled $\bot$, which indicates that there are not enough queries inside the ball to determine it's label.

@@

The simplified algorithm addresses the first challenge of ``where
to query'' (see challenges in the introduction). It does not address
the other two challenges, which the detailed algorithm does address.

@@

A related problem is that having the outer loop iterate over levels will
cause the algorithm to miss uncertain regions that later appear in lower
levels as a result of background sampling (the so-called mind changes).
Instead, each iteration of the full algorithm looks for the lowest level 
in which the uncertainty region is non-empty, and makes just two queries, 
one focused and one background.

@@

We compare two algorithms, the ``active'' agorithm described here, and the ``passive'' algorithm which uses only background queries. We denote the difference between the error of the generated classifier and that of the Bayes optimal classifier by $\epsilon$. From the theory, we expect the query complexity of the passive algorithm to be $O(\frac{1}{\epsilon})$ and of the active algorithm to be $O(\log (\frac{1}{\epsilon}))$.

>>>>

\usepackage{titlesec}
\titlespacing*{\section}{0pt}{0pt}{0pt}
\titlespacing*{\subsection}{0pt}{0pt}{0pt}
\titlespacing*{\subsubsection}{0pt}{0pt}{0pt}

@@

\begin{abstract}%
  We present a general-purpose active learning scheme that maintains a
  collection of balls of different sizes and uses label
  queries to identify those with a strong bias towards a particular
  label. When two such balls intersect and have different
  labels, the region of overlap is treated as a ``known unknown'' and
  is targeted in future active queries. We give label complexity
  bounds for this method that do not rely on any assumptions about the
  data, and we instantiate them in several cases of interest.
\end{abstract}

@@

As is well known~\cite{CH67, S80}, the $k$-NN method converges to
Bayes optimal rule without any assumptions on the underlying
distribution. Convergence rate is not uniform. It is slower at points
close to the decision boundary and faster at points further from the
boundary.  Our goal is to mimic the asymptotic performance of $k$-NN
while achieving lower query complexity by reducing the number of
queries far from the boundary.

We begin with a large collection of \emph{balls}, $\B$, of varying
sizes (Fig.~\ref{fig:challenges}(a)). We use label queries to assign
\emph{biases} $\yh(B) \in \{-1,0,+1\}$ to balls $B \in \B$, beginning
with the largest balls. A bias of $+1$ means that the average of 
$\eta(x)$ over $x \in B$ is significantly positive; bias $-1$ means it is
significantly negative; and bias $0$ means the average $\eta$ is close
to 0. As time goes on, we get biases for progressively smaller balls,
as needed. The label $\yh(x)$ assigned to any point $x \in X$ is based on the
biases of balls containing $x$.

@@

The rest of the paper is organized as
follows. Section~\ref{sec:relatedwork} describes related
work. Section~\ref{sec:definitions} defines some terminology and
notation.  Section~\ref{sec:algorithm} presents the active learning
algorithm and Section~\ref{sec:discrete} analyzes it in the finite
population setting, taking $X$ to be an arbitrary point-set and
allowing any $\eta$ function. We identify two \emph{critical levels}
for any $x \in X$: two scales (of ball sizes) that control how many
queries are sufficient for $x$ to be correctly labeled
(Theorem~\ref{thm:label-complexity}). We instantiate these bounds in
canonical one-dimensional settings (Theorems~\ref{thm:oned-massart}
and \ref{thm:oned-monotonic}) to get label complexities logarithmic in
$|X|$. In section~\ref{sec:experiment} we describe an experiment that
supports the theoretical bound given in
Theorem~\ref{thm:oned-massart}.

@@

\section{Definitions}
\label{sec:definitions}
The two basic entities in our algorithm are the collection of points
$X$ and a collection of balls $\B$. Sampling is organized around $\B$:
these balls are the atomic sets on which we assess label
bias.

@@

We now describe the simplified algorithm. The algorithm loops over
increasing levels.  At each level it identifies the region of ``known
unknowns'' which approximates the boundary. The algorithm then makes
queries in the vicinity of the boundary to refine it and proceeds to
the next finer level.

Figure~\ref{alg:simple} contains the pseudo-code of the simplified algorithm.
What follows is a detailed explanation of this code.
The outer loop of the algorithm iterates over
levels, starting with $\ell=0$ that corresponds to balls with at least half of the points in $X$, then
$\ell=1$ which has balls with between $1/4$ and $1/2$ of the points in $X$, and so on.

@@

The simplified algorithm addresses the first challenge of ``where
to query'' (see challenges in the introduction). It does not address
the other two challenges, addressed here with the detailed algorithm.

@@

A related problem is that having the outer loop iterate over levels might
cause the algorithm to miss uncertain regions that later appear in lower
levels as a result of background sampling (the so-called mind changes).
Instead, each iteration of the detailed algorithm looks for the lowest level 
in which the uncertainty region is non-empty, and makes just two queries, 
one focused and one background.

@@

We compare two algorithms, the ``active'' algorithm described here, and the ``passive'' algorithm which uses only background queries. We denote the difference between the error of the generated classifier and that of the Bayes optimal classifier by $\epsilon$. From the theory, we expect the query complexity of the passive algorithm to be $O(\frac{1}{\epsilon})$ and of the active algorithm to be $O(\log (\frac{1}{\epsilon}))$.
